### 2024-05-20 plots/combined/aft_comparison.tex
clean bibliography

\begin{table*}
\centering
\caption{Comparison of AFR Models on the Combined dataset. The metrics are defined in Sec.~\ref{afr}}
\label{tab:combined}
\begin{tabular}{lrrrrrr}
\toprule
 & AIC & BIC & Concordance & Test Concordance & ICI & E50 \\
\midrule
Weibull & -5.57e+03 & -5.57e+03 & 0.84 & 0.84 & 0 & 0 \\
Log Logistic & -5.71e+03 & -5.71e+03 & 0.85 & 0.84 & 0.01 & 0 \\
Log Normal & -5.85e+03 & -5.85e+03 & 0.84 & 0.84 & 0.01 & 0 \\
\bottomrule
\end{tabular}
\end{table*}