### 2024-05-20 plots/cifar100/aft_comparison.tex

\begin{table*}
\centering
\caption{Comparison of AFR Models on the CIFAR100 dataset.}
\label{tab:cifar100}
\begin{tabular}{lllrrrrrr}
\toprule
 & AIC & BIC & Concordance & Test Concordance & ICI & Test ICI & E50 & Test E50 \\
\midrule
Weibull & -4.19e+03 & -4.19e+03 & 0.77 & 0.75 & 0 & 0 & 0 & 0 \\
Cox & -- & -- & 0.77 & 0.76 & 0.01 & 0.01 & 0 & 0 \\
Log Logistic & -4.49e+03 & -4.49e+03 & 0.77 & 0.75 & 0.02 & 0.01 & 0 & 0 \\
Log Normal & -4.65e+03 & -4.65e+03 & 0.77 & 0.75 & 0 & 0 & 0 & 0 \\
\bottomrule
\end{tabular}
\end{table*}