### 2024-08-20 plots/cifar100/aft_comparison.tex

\begin{table*}
\centering
\caption{Comparison of AFT Models on the CIFAR100 dataset.}
\label{tab:cifar100}
\begin{tabular}{lllrrrrrr}
\toprule
 & AIC & BIC & Concordance & Test Concordance & ICI & Test ICI & E50 & Test E50 \\
\midrule
Weibull & -4.4e+03 & -4.4e+03 & 0.81 & 0.8 & 0 & 0.01 & 0 & 0 \\
Cox & -- & -- & 0.81 & 0.8 & 0.01 & 0.01 & 0 & 0 \\
Log Logistic & -4.73e+03 & -4.73e+03 & 0.82 & 0.81 & 0.01 & 0.02 & 0 & 0 \\
Log Normal & -4.89e+03 & -4.89e+03 & 0.82 & 0.81 & 0 & 0.02 & 0 & 0 \\
\bottomrule
\end{tabular}
\end{table*}