### 2024-08-22 glossary.tex

\newacronym{aft}{AFT}{Accelerated Failure Time}
\newacronym{cpu}{CPU}{central processing unit}
\newacronym{e50}{E50}{error at 50th percentile}
\newacronym{fgm}{FGM}{Fast Gradient Method}
\newacronym{aic}{AIC}{Akaike Informaiton Criterion}
\newacronym{gpu}{GPU}{graphics processing unit}
\newacronym{ici}{ICI}{integrated calibration index}
\newacronym{bic}{BIC}{Bayesian Information Criteria}
\newglossaryentry{kepler}
{
name=KEPLER,
description={Kubernetes Efficient Power Level Exporter}
}
\newacronym{ml}{ML}{machine learning}

% GB
% VRAM
% MNIST
% CIFAR
% CIFAR100
% GCP