### 2024-09-11 cover_letter.tex
Updates from Overleaf

\documentclass{letter}

\signature{Charles Meyers, \\ Mohammed Reza, \\ Tommy L\"{o}fstedt, \\ and  Erik Elmroth \\ }
\address{Department of \\ Computing Science,\\ Ume\aa University,\\ Sweden}

\begin{document}
\begin{letter}{}
\opening{To whom it may concern}

I am writing you on behalf of the authors of \em{A Cost-Aware Approach to Adversarial Robustness in Neural Networks}, which is the attached manuscript.  In this work, we developed a cloud-native methodology for quickly and precisely modelling the effect of model covariates using accelerated failure time models and a multi-objective search to minimize training cost while maximizing model accuracy. Not only do we provide a detailed outline and related source code, but we show the surprising result that smaller, more power-efficient GPUs create models that are more robust and more cost effective than models built on hardware that is advertised as being more capable. 
\closing{Sincerely,}
\end{letter}
\end{document}