### 2024-09-11 cover_letter.tex
Updates from Overleaf

\documentclass{letter}

\signature{Charles Meyers}
\address{Department of \\ Computing Science,\\ Ume\aa{}~University,\\ Sweden}

\begin{document}
\begin{letter}{}
\opening{Dear Editors,}

I am writing you on behalf of the authors of \textit{A Cost-Aware Approach to Adversarial Robustness in Neural Networks}, which is the attached manuscript.  In this work, we developed a cloud-native methodology for quickly and precisely modelling the effect of model covariates using accelerated failure time models and a multi-objective search to minimize training cost while maximizing model accuracy. Not only do we provide a detailed outline and related source code, but we show the surprising result that smaller, more power-efficient GPUs create models that are more robust and more cost effective than models built on hardware that is advertised as being more capable.

% You can add something about why we submit to this journal (showing that we are in scope and that we think highly of the journal) and that we look forward to a swift processing of the manuscript.)

\closing{On behalf of the authors, sincerely,}
\end{letter}
\end{document}