### 2024-09-11 plots/cifar/aft_comparison.tex
update plots

\begin{table*}
\centering
\caption{Comparison of AFT Models on the CIFAR dataset.}
\label{tab:cifar}
\begin{tabular}{lllrrrrrr}
\toprule
 & AIC & BIC & Concordance & Test Concordance & ICI & Test ICI & E50 & Test E50 \\
\midrule
Weibull & -5.18e+03 & -5.19e+03 & 0.85 & 0.86 & 0 & 0.01 & 0 & 0 \\
Cox & -- & -- & 0.85 & 0.86 & 0.02 & 0.02 & 0 & 0 \\
Log Logistic & -5.42e+03 & -5.43e+03 & 0.84 & 0.85 & 0.01 & 0.01 & 0 & 0 \\
Log Normal & -5.57e+03 & -5.58e+03 & 0.84 & 0.85 & 0.01 & 0.01 & 0 & 0 \\
\bottomrule
\end{tabular}
\end{table*}