### 2024-09-11 plots/mnist/aft_comparison.tex

\begin{table*}
\centering
\caption{Comparison of AFT Models on the MNIST dataset.}
\label{tab:mnist}
\begin{tabular}{lllrrrrrr}
\toprule
 & AIC & BIC & Concordance & Test Concordance & ICI & Test ICI & E50 & Test E50 \\
\midrule
Weibull & -6.46e+03 & -6.47e+03 & 0.9 & 0.9 & 0.01 & 0.01 & 0 & 0 \\
Cox & -- & -- & 0.9 & 0.9 & 0.03 & 0.03 & 0 & 0 \\
Log Logistic & -6.47e+03 & -6.48e+03 & 0.91 & 0.91 & 0.04 & 0.04 & 0 & 0 \\
Log Normal & -6.61e+03 & -6.62e+03 & 0.9 & 0.91 & 0.03 & 0.03 & 0 & 0 \\
\bottomrule
\end{tabular}
\end{table*}